---
title: "Basic Needs Center Peer Conversations"
author: "Noah Adams"
format: html
editor: visual
---

```{r}
#| echo: false
suppressPackageStartupMessages(library(tidyverse))
```

```{r}
#| echo: false
pal <- c("#6C88C4","#FF828B","#E7C582")
```

```{r}
#| echo: false
referrals <- read.csv('Referrals.csv')

month_order <- c("Sept", "Oct", "Nov", "Dec", "Jan", "Feb","Mar")

referrals <- referrals |>
  pivot_longer(cols = -1,
                names_to = 'Program',
                values_to = 'Count') |>
    rename(Month = Referral) |>
  mutate(Month = factor(Month, levels = month_order))
```

# What is a Peer Conversation?

At the OSU Basic Needs Center, we have strategy conversations with students about housing, food, living on a limited budget, and anything else related to someone’s basic needs. Following these conversations, we fill out a form recording what referrals were made, and what the conversation was about. In this report, I hope to analyze and further explore the peer conversations we have at the BNC.

*Please note that the data for September has been measured only from the 15th to the 30th.*

## Conversations Over Time

```{r}
#| label: Monthly Peer Convos
#| echo: false

convo_counts = c(77, 51, 37, 17, 56, 23, 40)

convo_totals <- data.frame(Month = factor(month_order, levels = month_order), Count = convo_counts)

ggplot(convo_totals, aes(x = Month, y = Count)) +
  geom_col() +
  theme(text=element_text(size=13, family="Arial")) +
  geom_text(aes(label = Count), vjust = 2, color = "white", size = 3.5) +
  theme_bw() +
  labs(
    title = 'Monthly Peer Convos'
  )
```

## Referrals During Peer Conversations

### Summary of Referrals

```{r}
#| label: Summary of Referrals
#| echo: false

referral_summary <- referrals |>
  group_by(Program) |>
  summarize(Count = sum(Count)) |>
  arrange(desc(Count)) |>
  mutate(Program = factor(Program, levels = Program))

programs = c('BNC Program', 'BNC Food Pantry', 'SNAP/WIC', 'Other College Office', 'Housing Org or Shelter', 'Other Community Org', 'Financial Aid/CAFE/Student Account', 'Off-campus Food Pantry', 'OHP', 'Career Development Ctr', 'Social Services', 'Career/Employment Org', 'Legal Support', 'Housing Authority', 'Childcare', 'Other Govt Agency', 'Student Care', 'TANF')

ggplot(referral_summary, aes(x = Program, y = Count)) +
  geom_col() +
  scale_x_discrete(labels = programs) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text=element_text(size=13, family="Arial")) +
  labs(
    title = 'Program Referrals'
  )
```

## Subsets of Program Data

### Food Program Referrals

```{r}
#| label: Food Program Referrals
#| echo: false

# Isolate the food programs
food_programs <- referrals |>
  filter(Program == c('SNAP.WIC', 'BNC.Food.Pantry', 'Off.Campus.Food.Bank')) |>
  mutate(Program = factor(Program, levels = c('BNC.Food.Pantry', 'SNAP.WIC', 'Off.Campus.Food.Bank')))
  
# Create barplot of food programs
ggplot(data = food_programs, aes(x = Month, y = Count, fill = Program)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(text=element_text(size=13, family="Arial")) +
  labs(
    title = 'Food Program Referrals'
  )
```

### Housing Program Referrals

```{r}
#| label: Housing Program Referrals
#| echo: false

# Isolate housing programs
housing_programs <- referrals |>
  filter(Program == 'Housing.Org..or.Shelter' | Program == 'Housing.Authority')

ggplot(data = housing_programs, aes(x = Month, y = Count, fill = Program)) +
  geom_col() +
  theme_bw() +
  theme(text=element_text(size=13, family="Arial")) +
  labs(
    title = 'Housing Program Referrals'
  )
```

### Heatmap of Referrals

```{r}
#| echo: false

heatmap_programs <- c('BNC Food Pantry', 'BNC Program/Service', 'Career/Employment Org', 'Career Development Ctr', 'Childcare', 'Financial Aid')

ggplot(referrals, aes(Month, Program)) +
  geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "#E15759") +
  theme_bw() +
  scale_y_discrete() +
  theme(text=element_text(size=13, family="Arial"))
```

## Topics of Conversations

```{r}
#| echo: false
topics <- read.csv('Topics.csv')

topics <- topics |>
  pivot_longer(cols = -1,
                names_to = 'Program',
                values_to = 'Count') |>
    rename(Month = Topic) |>
    mutate(Month = factor(Month, levels = month_order))

```

### Summary of Topics

```{r}
#| label: Summary of Topics
#| echo: false

topics_summary <- topics |>
  group_by(Program) |>
  summarize(Count = sum(Count)) |>
  arrange(desc(Count)) |>
  mutate(Program = factor(Program, levels = Program))

ggplot(topics_summary, aes(x = Program, y = Count)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text=element_text(size=13, family="Arial")) +
  labs(
    title = 'Summary of Convo Topics'
  )
```

### Three Main Topics

```{r}
#| label: Three Main Topics
#| echo: false

top3 <- topics |>
  filter(Program == 'Food' | Program == 'Housing' | Program == 'Technology')

ggplot(top3, aes(x = Month, y = Count, fill = Program)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Program) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    text=element_text(size=11, family="Arial")) +
  labs(
    title = 'Three Main Topics'
  )
```

### Bubble Chart of Topics

```{r}
#| echo: false
ggplot(topics, aes(x = Month, y = Program, size = Count)) +
  geom_point(alpha = 0.7) +
  theme_bw() +
  theme(text=element_text(size=13, family="Arial"))
```
