---
title: "BNC Peer Convos"
subtitle: "A graphical analysis of peer convos from the year 2025-2026"
author: "Noah Adams"
format: revealjs
editor: visual
---

```{r}
#| echo: false
suppressPackageStartupMessages(library(tidyverse))
```

```{r}
#| echo: false
pal <- c("#6C88C4","#FF828B","#FFB347")
```

```{r}
#| echo: false

referrals <- read.csv('Referrals.csv')

month_order <- c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr")

referrals <- referrals |>
  pivot_longer(cols = -1,
                names_to = 'Program',
                values_to = 'Count') |>
    rename(Month = Referral) |>
  mutate(Month = factor(Month, levels = month_order))
```

## What is a Peer Convo?

- At the OSU Basic Needs Center, we have strategy conversations with students about food, housing, living on a limited budget, and anything else related to someone’s basic needs.

- Following these conversations, we fill out a form recording what referrals were made, and what the conversation was about.

- In this presentation, I hope to analyze and further explore the peer conversations we have at the BNC.

## Convos This Year

A noticeable pattern is the big surges of convos at the start of the term. The drop in December could be explained by Winter Break.

```{r}
#| label: Monthly Peer Convos
#| echo: false

convo_counts = c(77, 51, 37, 17, 56, 23, 40, 44)

convo_totals <- data.frame(Month = factor(month_order, levels = month_order), Count = convo_counts)

ggplot(convo_totals, aes(x = Month, y = Count)) +
  geom_col() +
  theme(text=element_text(size=13, family="sans")) +
  geom_text(aes(label = Count), vjust = 2, color = "white", size = 3.5) +
  theme_bw() +
  labs(
    title = 'Monthly Peer Convos'
  )
```
