---
title: "Basic Needs Center: Peer Conversations"
author: "Noah Adams"
date: today
format: html
editor: visual

include-in-header:
  - text: |
     <link rel = "shortcut icon" href = "logo.png" />
---

```{r}
#| label: Libraries
#| echo: false
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gt))
```

```{r}
#| label: Palette
#| echo: false
pal <- c("#6C88C4","#FF828B","#FFB347")
```

```{r}
#| label: Referral Data Processing
#| echo: false
referrals <- read.csv('Referrals.csv')

month_order <- c("Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr")

referrals <- referrals |>
  pivot_longer(cols = -1,
                names_to = 'Program',
                values_to = 'Count') |>
    rename(Month = Referral) |>
  mutate(Month = factor(Month, levels = month_order)) |>
  mutate(Program = recode(Program,
  'BNC.Food.Pantry' = 'BNC Food Pantry',
  'SNAP.WIC' = 'SNAP/WIC',
  'Off.Campus.Food.Bank' = 'Off Campus Food Bank',
  'BNC.Program.Service' = 'BNC Program',
  'Other.College.Dept..Office' = 'Other College Dept or Office',
  'Housing.Org..or.Shelter' = 'Housing Org or Shelter',
  'Other.Community.Based.Org.' = 'Other Community Based Org',
  'Financial.Aid..CAFE..Student.Accounts..Emergency.grants' = 'Financial Aid/CAFE/Emergency Grants',
  'Oregon.Health.Plan..OHP.' = 'Oregon Health Plan (OHP)',
  'Career.Development.Ctr' = 'Career Development Ctr',
  'Social.Services..Self.Help' = 'Social Services / Self Help',
  'Career...Employment.Org.' = 'Career Employment Org',
  'Housing.Authority' = 'Housing Authority',
  'Legal.Support' = 'Legal Support',
  'Other.Govt..Agency' = 'Other Govt Agency',
  'Child.Care' = 'Child Care',
  'Student.Care' = 'Student Care'))
```

# What is a Peer Conversation?

At the OSU Basic Needs Center, we have strategy conversations with students about housing, food, living on a limited budget, and anything else related to someone’s basic needs. Following these conversations, we fill out a form recording what referrals were made, and what the conversation was about. In this report, I hope to analyze and further explore the peer conversations we have at the BNC.

*Please note that the data for September has been measured only from the 15th to the 30th.*

## Conversations Over Time

A noticeable pattern is the big surges of convos at the start of the term. The drop in December could be explained by Winter Break.

```{r}
#| label: Monthly Peer Convos
#| echo: false

convo_counts = c(77, 51, 37, 17, 56, 23, 40, 44)

convo_totals <- data.frame(Month = factor(month_order, levels = month_order), Count = convo_counts)

ggplot(convo_totals, aes(x = Month, y = Count)) +
  geom_col() +
  theme(text=element_text(size=13, family="sans")) +
  geom_text(aes(label = Count), vjust = 2, color = "white", size = 3.5) +
  theme_bw() +
  labs(
    title = 'High number of convos at the start of the term'
  )
```

## Referrals During Peer Conversations

Frequently during convos, we refer people to other resources both on and off campus. Helping people utilize the full network of resources available to them is crucial.

### Summary of Referrals

```{r}
#| label: Summary of Referrals
#| echo: false

referral_summary <- referrals |>
  group_by(Program) |>
  summarize(Count = sum(Count)) |>
  arrange(desc(Count)) |>
  mutate(Program = factor(Program, levels = Program))

ggplot(referral_summary, aes(x = Program, y = Count)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text=element_text(size=13, family="sans")) +
  labs(
    title = 'BNC Programs are leading referrals'
  )
```

### Subsets of Program Data

Below, the referrals made for resources relevant to two of our major areas of support, food and housing, are isolated and count is shown over time.

#### Food Program Referrals

```{r}
#| label: Food Program Referrals
#| echo: false

# Isolate the food programs
food_programs <- referrals |>
  filter(Program == c('SNAP/WIC', 'BNC Food Pantry', 'Off Campus Food Bank')) |>
  mutate(Program = factor(Program, levels = c('BNC Food Pantry', 'SNAP/WIC', 'Off Campus Food Bank')))

# Create barplot of food programs
ggplot(data = food_programs, aes(x = Month, y = Count, fill = Program)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(text=element_text(size=13, family="sans")) +
  labs(
    title = 'Food Program Referrals per Month'
  )
```

While referrals to the BNC Food Pantry are high, the bar chart above and the table below highlight that we don't make many referrals to off-campus food banks, when maybe we should as part of someone's food strategy.

```{r}
#| label: Food Pantry Referrals Table
#| echo: false

food_pantries <- referrals |>
  filter(Program == c('Off Campus Food Bank', 'BNC Food Pantry')) |>
  mutate(Program = factor(Program, levels = c('BNC Food Pantry', 'Off Campus Food Bank')))

pantries_wide <- food_pantries |>
  pivot_wider(names_from = Program, values_from = Count)

pantries_wide |>
  gt() |>
  tab_header(
    title = 'Food Pantry Referrals',
    subtitle = 'BNC Food Pantry has significantly higher referrals than off campus'
  ) |>
  tab_options(
    heading.background.color = '#dcd5e7'
  ) |>
  opt_table_font(
    font = 'sans'
  )
```

#### Housing Program Referrals

```{r}
#| label: Housing Program Referrals
#| echo: false

# Isolate housing programs
housing_programs <- referrals |>
  filter(Program == 'Housing Org or Shelter' | Program == 'Housing Authority')

ggplot(data = housing_programs, aes(x = Month, y = Count, fill = Program)) +
  geom_col(position = 'dodge') +
  theme_bw() +
  theme(text=element_text(size=13, family="sans")) +
  labs(
    title = 'Housing Program Referrals'
  )
```

## Topics of Conversations

```{r}
#| echo: false
topics <- read.csv('Topics.csv')

topics <- topics |>
  pivot_longer(cols = -1,
                names_to = 'Program',
                values_to = 'Count') |>
    rename(Month = Topic) |>
    mutate(Month = factor(Month, levels = month_order)) |>
  mutate(Program = recode(Program,
  'Academic.Resources...Support' = 'Academic Resources and Support',
  'Emergency.Finances' = 'Emergency Finances',
  'Career...Employment' = 'Career Employment',
  'Personal.Hygiene.Items' = 'Personal Hygiene Items',
  'Legal.Services' = 'Legal Services',
  'Other.need.disclosed' = 'Other Need Disclosed',
  'Physical.Health.Care' = 'Physical Health Care',
  'Mental.Health.Care' = 'Mental Health Care',
  'Disability.Services' = 'Disability Services'
  ))

```

### Summary of Topics

Breaking down what kinds of conversations we are having, a clear trend arises of focusing on food, housing, and technology.

```{r}
#| label: Summary of Topics
#| echo: false

topics_summary <- topics |>
  group_by(Program) |>
  summarize(Count = sum(Count)) |>
  arrange(desc(Count)) |>
  mutate(Program = factor(Program, levels = Program))

ggplot(topics_summary, aes(x = Program, y = Count)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text=element_text(size=13, family="sans")) +
  labs(
    title = 'Food, housing, and technology are main convo topics'
  )
```

### Three Main Topics

```{r}
#| echo: false

top3 <- topics |>
  filter(Program == 'Food' | Program == 'Housing' | Program == 'Technology')

ggplot(top3, aes(x = Month, y = Count, group = Program)) +
  geom_line(aes(color = Program)) +
  geom_point(aes(color = Program)) +
  scale_color_manual(values = pal) +
  theme_bw() +
  labs(
    title = 'Food, housing, and technology trends over time'
  )
```

## Acronyms

`BNC`: Basic Needs Center

`SNAP`: Supplemental Nutrition Assistance Program

`WIC`: Women, Infants, and Children (nutrition assistance)

`CAFE`: Center for Advancing Financial Education

`OHP`: Oregon Health Plan

`TANF`: Temporary Assistance for Needy Families (cash benefits)

```{r}
#| label: Bubble Plot
#| eval: false
#| echo: false
ggplot(topics, aes(x = Month, y = Program, size = Count)) +
  geom_point(alpha = 0.7) +
  theme_bw() +
  theme(text=element_text(size=13, family="sans"))
```

```{r}
#| label: Heatmap
#| eval: false
#| echo: false

ggplot(referrals, aes(Month, Program)) +
  geom_tile(aes(fill = Count), color = "white") +
  scale_fill_gradient(low = "white", high = "maroon") +
  theme_bw() +
  scale_y_discrete() +
  theme(text=element_text(size=13, family="sans"))
```

```{r}
#| label: BNC vs. Outside Pantry Referrals
#| echo: false
#| eval: false

food_pantries <- referrals |>
  filter(Program == c('Off.Campus.Food.Bank', 'BNC.Food.Pantry')) |>
  mutate(Program = factor(Program, levels = c('BNC.Food.Pantry', 'Off.Campus.Food.Bank')))

labels <- c('BNC Food Pantry', 'Off Campus Food Bank')

ggplot(food_pantries, aes(Program, Count, color = Program)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.09) +
  theme_bw() +
  theme(text=element_text(size=13, family="sans"),
        legend.position = "none")+
  labs(
    title = 'Monthly Food Program Referrals',
  ) +
  scale_x_discrete(label = labels)
```

```{r}
#| label: Three Main Conversation Topics
#| echo: false
#| eval: false

top3 <- topics |>
  filter(Program == 'Food' | Program == 'Housing' | Program == 'Technology')

ggplot(top3, aes(x = Month, y = Count, fill = Program)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Program) +
  scale_fill_manual(values = pal) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    text=element_text(size=11, family="sans")) +
  labs(
    title = 'Three Main Conversation Topics'
  )
```
